Parsing Postfix log files

### 2008-04-09 doc/logparser.tex
Reformatted other parsers section. Added TODO items.

XXX POSSIBLE CONFERENCES\@: SIGAI IN BRITAIN, AND AICS'S AI-2008 (CAMBRIDGE
IN DECEMBER)\@.

@@

@@

    \item Only \parsername{} makes it possible to parse new log lines
        without modifying the core parsing algorithm; see \sectionref{why
        separate rules, actions and framework?} for a discussion of why the
        separation of rules from the algorithm is beneficial.

    \item \parsername{} does not provide any reporting based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        logs and extracting data --- further processing of that data is a
        separate concern, deferred to another program.  Using an \SQL{}
        database simplifies the process of generating such reports; some
        simple examples are given in \sectionref{motivation}.

@@

In this section text in italics is a quote extracted from the appropriate
program's documentation.

    \item [Pflogsumm] \textit{pflogsumm is designed to provide an over-view
        of Postfix activity, with just enough detail to give the
        administrator a ``heads up'' for potential trouble spots.\/}

@@

    \item [Sawmill Universal Log File Analysis and Reporting] \textit{
        Sawmill is a \newline Postfix log analyzer (it also support 686
        other log formats).  It can process log files in Postfix format,
        and generate dynamic statistics from them, analyzing and reporting
        events.  Sawmill can parse Postfix logs, import them into a SQL
        database (or its own built-in database), aggregate them, and
        generate dynamically filtered reports, all through a web interface.
        Sawmill can perform Postfix analysis on any platform, including
        Window, Linux, FreeBSD, OpenBSD, Mac OS, Solaris, other UNIX, and
        more.\/}

@@

    \item [Splunk] \textit{Splunk is an IT Search engine. It is software
        that indexes any format of IT data from any source in real time,
        including logs, configurations, scripts, code, messages, traps,
        alerts, activity reports, stack traces and metrics from all of your
        applications, servers and devices. Splunk lets you search,
        navigate, alert and report on all your IT data in real time using
        an AJAX web interface. You can also share knowledge and Splunk
        solutions with other members of the Splunk community via
        SplunkBase.\/}

@@

    \item [Isoqlog] \textit{Isoqlog is an MTA log analysis program written
        in C. It designed to scan qmail, postfix, sendmail and exim logfile
        and produce usage statistics in \HTML{} format for viewing through
        a browser. It produces Top domains output according to Sender,
        Receiver, Total mails and bytes; it keeps your main domain mail
        statistics with regard to Days Top Domain, Top Users values for per
        day, per month and years.\/}

@@

    \item [AWStats] \textit{AWStats is a free powerful and featureful tool
        that generates advanced web, streaming, ftp or mail server
        statistics, graphically. This log analyzer works as a CGI or from
        command line and shows you all possible information your log
        contains, in few graphical web pages. It uses a partial information
        file to be able to process large log files, often and quickly. It
        can analyze log files from all major server tools like Apache log
        files (NCSA combined/XLF/ELF log format or common/CLF log format),
        WebStar, IIS (W3C log format) and a lot of other web, proxy, wap,
        streaming servers, mail servers and some ftp servers.\/}

@@

        \url{http://awstats.sourceforge.net/} \newline
        \url{http://awstats.sourceforge.net/awstats.mail.html} \newline
        \url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL}
        \newline (Last checked 2007/10/13.)

    \item [Log analyser --- throughput monitor] This utility tracks the
        number of events which occurred over a particular time and warns if
        the frequency of events passes a certain threshold.  It's designed
        to provide real time alerts when dictionary attacks, mail loops or
        similar problems occur. It doesn't attempt to extract or save data,
        correlate log lines, or any of the more advanced tasks described in
        this document, because it is not designed to do so.  The user must
        construct the \regexes{} to match significant input lines before it
        will parse any logs.  This program was not reviewed in further
        detail because its aims are so far removed from the aims of this
        project.

@@

    \item [Anteater] \textit{The Anteater project is a Mail Traffic
        Analyser. Anteater supports currently the logformat produced by
        Sendmail and by Postfix. The tool is written in 100\% C++ and is
        very easy to customize. Input, output, and the analysis are modular
        class objects with a clear interface. There are eight useful
        analyse modules, writing the result in plain ASCII or \HTML{}, to
        stdout or to files.\/}

@@

    \item [Yet Another Advanced Logfile Analyser] \textit{yaala is a very
        flexible analyser for all kinds of logfiles. It uses parsers to
        extract information from a logfile, an SQL-like query language to
        relate the information to each other and an output-module to format
        the information appropriately.\/}

@@

        \tab{}Aggregations: count (not explained), bytes (sum of bytes
        transferred).\newline \tab{}Keyfields [sic]: date, hour, sender,
        recipient, defer\_count, delay, incoming\_host, outgoing\_host.

        YAALA extracts most of the fields \parsername{} does, but it does
        not maintain separate counters for each restriction like
        \parsername{}, so its data cannot be used for optimisation, testing
        or understanding of restrictions.  YAALA's Postfix parser does not
        deal with the complications explained in this document, though it
        does correlate log lines by queueid.

@@

    \item [Logparser/Lire] \textit{As any good system administrator knows,
        there's a lot more to keep track of in an active network than just
        webservers. Lire is hands down the most versatile log analysis
        software available today. Lire not only keeps you informed about
        your HTTP, FTP, and mail traffic, it also reports on your
        firewalls, your print servers, and your DNS activity. The ever
        growing list of Lire-supported services clearly outstrips any other
        software, in large part thanks to the numerous volunteers who have
        pioneered many new services and features. Lire is a total solution
        for your log analysis needs.\/}

@@

    \item [Logrep] \textit{Logrep is a secure multi-platform framework for
        the collection, extraction, and presentation of information from
        various log files. It features HTML reports, multi dimensional
        analysis, overview pages, SSH communication, and graphs, and
        supports over 30 popular systems including Snort, Squid, Postfix,
        Apache, Sendmail, syslog, ipchains, iptables, NT event logs,
        Firewall-1, wtmp, xferlog, Oracle listener and Pix.\/}

@@

    \item [Log Mail Analyser] Please see the previous in-depth discussion
        of Log Mail Analyser in \sectionref{prior art}.

>>>>

XXX POSSIBLE CONFERENCES\@: AICS'S AI-2008 IN CORK, AND SGAI IN CAMBRIDGE
IN DECEMBER\@.

XXX DOWNLOAD\@: PFLOGSUMM, SAWMILL AND ANTEATER\@.

CHECK PARSING FOR THE FOLLOWING\@:

    \item Pflogsumm
    \item Sawmill Universal Log File Analysis and Reporting
    \item Splunk
    \item Isoqlog
    \item AWStats
    \item Log analyser --- throughput monitor
    \item Anteater
    \item Yet Another Advanced Logfile Analyser
    \item Logparser/Lire
    \item Logrep
    \item Log Mail Analyser

@@

\label{database as API}

@@

    \item Only \parsername{} makes it possible to parse new log lines
        without modifying the core parsing algorithm; see \sectionref{why
        separate rules, actions and framework?} for a discussion of why the
        separation of the algorithm into rules, actions and framework is
        beneficial.

    \item \parsername{} does not provide any reporting based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        logs and extracting data --- further processing of that data is a
        separate concern, deferred to another program.  Using an \SQL{}
        database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some
        simple examples are given in \sectionref{motivation}.

@@

In this section text in italics is a quote extracted from the documentation
of the program being reviewed.

\subsubsection{Pflogsumm}

\textit{pflogsumm is designed to provide an over-view of Postfix activity,
with just enough detail to give the administrator a ``heads up'' for
potential trouble spots.\/}

@@

\subsubsection{Sawmill Universal Log File Analysis and Reporting}

\textit{ Sawmill is a Postfix log analyzer (it also support 686 other log
formats).  It can process log files in Postfix format, and generate dynamic
statistics from them, analyzing and reporting events.  Sawmill can parse
Postfix logs, import them into a SQL database (or its own built-in
database), aggregate them, and generate dynamically filtered reports, all
through a web interface.  Sawmill can perform Postfix analysis on any
platform, including Window, Linux, FreeBSD, OpenBSD, Mac OS, Solaris, other
UNIX, and more.\/}

@@

\subsubsection{Splunk}

\textit{Splunk is an IT Search engine. It is software that indexes any
format of IT data from any source in real time, including logs,
configurations, scripts, code, messages, traps, alerts, activity reports,
stack traces and metrics from all of your applications, servers and
devices. Splunk lets you search, navigate, alert and report on all your IT
data in real time using an AJAX web interface. You can also share knowledge
and Splunk solutions with other members of the Splunk community via
SplunkBase.\/}

@@

\subsubsection{Isoqlog}

\textit{Isoqlog is an MTA log analysis program written in C. It designed to
scan qmail, postfix, sendmail and exim logfile and produce usage statistics
in \HTML{} format for viewing through a browser. It produces Top domains
output according to Sender, Receiver, Total mails and bytes; it keeps your
main domain mail statistics with regard to Days Top Domain, Top Users
values for per day, per month and years.\/}

@@

\subsubsection{AWStats}

\textit{AWStats is a free powerful and featureful tool that generates
advanced web, streaming, ftp or mail server statistics, graphically. This
log analyzer works as a CGI or from command line and shows you all possible
information your log contains, in few graphical web pages. It uses a
partial information file to be able to process large log files, often and
quickly. It can analyze log files from all major server tools like Apache
log files (NCSA combined/XLF/ELF log format or common/CLF log format),
WebStar, IIS (W3C log format) and a lot of other web, proxy, wap, streaming
servers, mail servers and some ftp servers.\/}

@@

\noindent\url{http://awstats.sourceforge.net/} \newline
\url{http://awstats.sourceforge.net/awstats.mail.html} \newline
\url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL}
\newline (Last checked 2007/10/13.)

\subsubsection{Log analyser --- throughput monitor}

This utility tracks the number of events which occurred over a particular
time and warns if the frequency of events passes a certain threshold.  It's
designed to provide real time alerts when dictionary attacks, mail loops or
similar problems occur. It doesn't attempt to extract or save data,
correlate log lines, or any of the more advanced tasks described in this
document, because it is not designed to do so.  The user must construct the
\regexes{} to match significant input lines before it will parse any logs.
This program was not reviewed in further detail because its aims are so far
removed from the aims of this project.

@@

\subsubsection{Anteater}

\textit{The Anteater project is a Mail Traffic Analyser. Anteater supports
currently the logformat produced by Sendmail and by Postfix. The tool is
written in 100\% C++ and is very easy to customize. Input, output, and the
analysis are modular class objects with a clear interface. There are eight
useful analyse modules, writing the result in plain ASCII or \HTML{}, to
stdout or to files.\/}

@@

\subsubsection{Yet Another Advanced Logfile Analyser}

\textit{yaala is a very flexible analyser for all kinds of logfiles. It
uses parsers to extract information from a logfile, an SQL-like query
language to relate the information to each other and an output-module to
format the information appropriately.\/}

@@

\noindent\tab{}Aggregations: count (not explained), bytes (sum of bytes
transferred).\newline \tab{}Keyfields [sic]: incoming\_host,
outgoing\_host, date, hour, sender, recipient, defer\_count, delay.

YAALA extracts most of the fields \parsername{} does, but it does not
maintain separate counters for each restriction like \parsername{}; this
rules out the possibility of using the collected data for  optimisation,
testing or understanding of restrictions.  YAALA's Postfix parser does not
deal with the complications explained in this document, though it does
correlate log lines by queueid.

@@

\subsubsection{Logparser/Lire}

\textit{As any good system administrator knows, there's a lot more to keep
track of in an active network than just webservers. Lire is hands down the
most versatile log analysis software available today. Lire not only keeps
you informed about your HTTP, FTP, and mail traffic, it also reports on
your firewalls, your print servers, and your DNS activity. The ever growing
list of Lire-supported services clearly outstrips any other software, in
large part thanks to the numerous volunteers who have pioneered many new
services and features. Lire is a total solution for your log analysis
needs.\/}

@@

\subsubsection{Logrep}

\textit{Logrep is a secure multi-platform framework for the collection,
extraction, and presentation of information from various log files. It
features HTML reports, multi dimensional analysis, overview pages, SSH
communication, and graphs, and supports over 30 popular systems including
Snort, Squid, Postfix, Apache, Sendmail, syslog, ipchains, iptables, NT
event logs, Firewall-1, wtmp, xferlog, Oracle listener and Pix.\/}

@@

\subsubsection{Log Mail Analyser}

Please see the previous in-depth discussion of Log Mail Analyser in
\sectionref{prior art}.
